### 2021-10-06 code/censusland_code.Rmd
Added census FIPS code + land cover data

---
title: "census-land-data"
output: html_document
---

##Load Dependencies
The following code loads the package dependencies for our analysis:
```{r}
#install.packages("tidycensus", repos = "http://cran.us.r-project.org")
#install.packages("censusapi", repos = "http://cran.us.r-project.org")
#install.packages("censusGeography", repos = "https://cran.rstudio.com/bin/macosx/contrib/4.0/chron_2.3-55.tgz")
#install.packages("rerddap", repos = "https://cran.rstudio.com/bin/macosx/contrib/4.0/rerddap_0.6.5.tgz")

library(tidyverse)
library(dplyr) #data wrangling
library(tidycensus) #access census data
library(censusapi)
library(rerddap)
library(RCurl)

#path to data
data_location <- here::here("data","processed_data","cleansampledata.rds")

#load data. 
data <- readRDS(data_location)
```

##Load Census API (load only once)
```{r}
#specify from which package a function is to be used

#tidycensus::census_api_key("c02ef0638e60f9d5fbb45de5219821e3a25cd7d2", install = TRUE) 
```
^This is not a code chunk you will need in every notebook. As long as "install = TRUE", you will only have to do this once. 
#restart R session after installing the first time

##Obtain FIPS codes for sampling sites
```{r}
#create function
latlong2fips <- function(latitude, longitude) {
  url <- "https://geo.fcc.gov/api/census/block/find?format=json&latitude=%f&longitude=%f"
  url <- sprintf(url, latitude, longitude)
  json <- RCurl::getURL(url)
  json <- RJSONIO::fromJSON(json)
  as.character(json$County['FIPS'])
}
#vectorize so function will take more than one argument
latlong2fips_vec <- Vectorize(latlong2fips, vectorize.args = "latitude", "longitude")
```

#accessing NHGIS land use/land cover data from downloaded file
source:https://www.nhgis.org/user-resources/environmental-summaries#download
```{r}
#path to data
data_location1 <- here::here("data","raw_data","landcover2015.csv")

#load data. 
landcover <- read.csv(data_location1)

class(landcover$GISJOIN)
#class = character, so I can use strsplit(). Data source uses extra letters and zeroes in their GIS code. I need to extract the original FIPS code to match my data

#extract state fip from GISJOIN
landcover = landcover %>% mutate(state_fip = substring(landcover$GISJOIN, 2,3)) %>% relocate(state_fip)

#extract county fip from GISJOIN
landcover = landcover %>% mutate(county_fip = substring(landcover$GISJOIN, 5, 7)) %>% relocate(county_fip)

#combine to state, county fip for whole fips codes
landcover$fips = paste0(landcover$state_fip, landcover$county_fip)

landcover <- landcover %>%
  select(fips, everything()) # moving fips column to far left of data set

#filter to clean data set
landcoverclean = landcover %>% select(-GISJOIN, -state_fip, -county_fip, -OID)

athensfips <- c("13219", "13059")
athensland <- filter(landcoverclean, fips %in% athensfips)

```