### 2021-10-26 analysis.Rmd
updated pop data, wrf data, land cover FIPS

---
title: "analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```