### 2021-10-27 analysis.Rmd
cleaned landcover data, began modeling analysis

---
title: "analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load packages and data
```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling

data <- readRDS(here("data", "processed_data", "athensland.rds"))
```

# fit linear model to continuous outcome (particles_l)
```{r}
#attempt to visualize data
#look at particles per L vs population
ggplot(data, aes(x = population, y = particles_l)) + geom_point() + geom_smooth(method = lm)
```
# linear regression
```{r}
#linear regression model specification set up
lm_mod <- linear_reg() %>% set_engine("lm")
#estimating/training linear model
lm_fit1 <- lm_mod %>% fit(particles_l ~ ., data = data)
#view summary of fit
tidy(lm_fit1)
```
# model evaluation
```{r}
#split into test and train
#set seed for reproducible analysis
set.seed(222)
#subset 3/4 of data as training set
data_split <- initial_split(data, prop = 3/4)

#save sets as data frames
train_data <- training(data_split)
test_data <- testing(data_split)
```

# recipe to fit continuous outcome to all predictors
```{r}
#recipe for particles per liter, all predictors
mp_rec <- 
  recipe(particles_l ~ ., data = train_data)

#linear regression model workflow set up
mp_wflow <- 
  workflow() %>% add_model(lm_mod) %>% add_recipe(mp_rec)

#use workflow to prepare recipe and train model with predictors
mp_fit <- 
  mp_wflow %>% fit(data = train_data)

#extract model coefficient
mp_fit %>% extract_fit_parsnip() %>% tidy()
```
# evaluate model with ROC and ROC-AUC
```{r}
#use trained workflow to predict with test data
test_data <- test_data[-c(20:28),]
predict(mp_fit, test_data)

#include probabilities
mp_aug <- augment(mp_fit, test_data)

#generate ROC curve
mp_aug %>% 
  roc_curve(truth = particles.l, .pred) %>% autoplot()

#estimate area under the curve
mp_aug %>% roc_auc(truth = particles_l, .pred)
```

Here, I am running into an issue since july_21 data is still missing. Hopefully we'll get that from Megan soon. For now, let's try filtering it out.