### 2021-11-12 products/manuscript/Supplement.Rmd

---
title: "Supplementary Materials: Plastic in the Urban Freshwater Environment"
subtitle: "An Exploratory Study of Microplastics in the Upper Oconee Watershed via Community-Based Efforts"
author: Amelia M. Foley
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
    global_numbering: true
    fig_caption: yes
    #reference_docx: style_template.docx #working on creating a template doc with the appropriate style for the conference I am attending next summer (ASTM) so that my manuscript can be automatically formatted
  bookdown::html_document2: 
    toc: false
bibliography: library.bib 
csl: elsevier.csl
---

## Exploratory analysis

This section will contain the key products of my exploratory analysis (located in exploration.Rmd) when complete. Since I am keeping this document in manuscript style, I will keep processing/exploration/analysis code separate.

Table \@ref(tab:summarytable) shows a table summarizing the data.
```{r summarytable,  echo=FALSE}

#summarytable=readRDS("../../results/summarytable.rds")
#knitr::kable(summarytable, caption = 'Data summary table.')

#I eventually want a nice looking summary statistic table. This is a place holder for now  
```

### Distribution of Microplastic Concentration

Figure \@ref(fig:distribution) shows a histogram of microplastic concentration observations. The minimum concentration is 16.67 particles/L and the maximum is 1193.33 particles/L. The mean concentration is 104.39 particles/L, and the median is 66.67 particles/L.

```{r distribution,  fig.cap='Distribution of Microplastic Concentration', echo=FALSE}
knitr::include_graphics("../../results/distribution.png")
```

Microplastic concentrations remained in similar ranges throughout the study period. Figure \@ref(fig:concentrationbydate) shows a boxplot of concentrations by sample date.

```{r concentrationbydate,  fig.cap='Particles/L by Sample Date', echo=FALSE}
knitr::include_graphics("../../results/concentrationbydate.png")
```

There is some seasonal variation in concentration at each individual site. Figure \@ref(fig:seasonalconc-bysite) shows a plot of concentrations at each site.

```{r seasonalconc-bysite,  fig.cap='Seasonal Variation in Particles/L', echo=FALSE}
knitr::include_graphics("../../results/seasonalconc_bysite.png")
```

There are similar microplastic levels throughout the watersheds within the Upper Oconee. Some watersheds experienced greater variation in microplastic levels than other watersheds. Figure \@ref(fig:conc-bywatershed) shows the microplastic concentrations by watershed.

```{r conc-bywatershed,  fig.cap='Watershed Microplastic Concentrations', echo=FALSE}
knitr::include_graphics("../../results/conc_bywatershed.png")
```

Figure \@ref(fig:watershed-overtime) shows a line graph of the mean watershed microplastic concentrations at each seasonal sampling date.

```{r watershed-overtime,  fig.cap='Microplastic Concentration Over Time', echo=FALSE}
knitr::include_graphics("../../results/watershed_overtime.png")
```

### Predictors
Population, land cover/use, and bacteria levels are hypothesized predictors of microplastic concentration. Figure \@ref(fig:particlesvpop) and Figure \@ref(fig:logmpvcfu) demonstrate the relationship between microplastic concentration and population and microplastic concentration and bacteria levels (CFU/L), respectively.

```{r particlesvpop,  fig.cap="Particles/L vs Population", echo=FALSE}
knitr::include_graphics("../../results/particlesvpop.png")
```

```{r logmpvcfu,  fig.cap='Log particles/L vs CFU', echo=FALSE}
knitr::include_graphics("../../results/logmpvcfu.png")
```

Figure \@ref(fig:matrix) and Figure \@ref(fig:landmatrix) show correlation matrices for the hypothesized predictor and for the different categories of land use.

```{r matrix,  fig.cap='Predictor matrix', echo=FALSE}
knitr::include_graphics("../../results/matrix.png")
```

```{r landmatrix,  fig.cap='Land cover matrix', echo=FALSE}
knitr::include_graphics("../../results/landmatrix.png")
```

## Full analysis

Preliminary modeling reveals that there is not a strong relationship between microplastic concentration and population level. Figure \@ref(fig:m-concvpop-no-outlier) demonstrates a linear model fit.

```{r m-concvpop-no-outlier,  fig.cap='Concentration vs Population Linear Model', echo=FALSE}
knitr::include_graphics("../../results/m_concvpop_no_outlier.png")
```

Figure \@ref(fig:m-logconcvcfu) shows a linear model of microplastic concentration vs CFU (both variables log-transformed).

```{r m-logconcvcfu,  fig.cap='Concentration vs CFU Linear Model', echo=FALSE}
knitr::include_graphics("../../results/m_logconcvcfu.png")
```

Figure \@ref(fig:concvturbidity) demonstrates a linear model of particles/L vs turbidity.

```{r m_concvturbidity,  fig.cap='Concentration vs Turbidity Linear Model', echo=FALSE}
knitr::include_graphics("../../results/m_concvturbidity.png")
```

Table \@ref(tab:lmfittable) shows a table summarizing a linear model fit predicting particles/L with 6 predictors.

```{r lmfittable,  echo=FALSE}
lm_fit_table=readRDS("../../results/lm_fit_table.rds")
knitr::kable(lm_fit_table, caption = 'Linear model fit table.')
```