### 2021-11-29 code/processing_code/map_images.Rmd
improve manuscript map

---
title: "map_images"
output: html_document
---

This document imports the map images available in results/maps and compiles them into one image containing 4 quads. The maps were generated by code on the author's local computer and then uploaded to this project repository due to issues with the mapping packages/files corrupting the GitHub/GitKraken interface.

```{r}
library(gridExtra)
library(png)
library(grid)
library(here)
library(tidyverse)

# map images for each sample date are saved in results/maps as 4 individual images
# here, we will combine the images in to a grid for use in the manuscript

map_fall <- readPNG(here("results", "maps", "fall_map.png"))
map_winter <- readPNG(here("results", "maps", "winter_map.png"))
map_spring <- readPNG(here("results", "maps", "spring_map.png"))
map_summer <- readPNG(here("results", "maps", "summer_map.png"))

map_fall

map_grid <- grid.arrange(rasterGrob(map_fall), rasterGrob(map_winter), rasterGrob(map_spring), rasterGrob(map_summer), ncol = 2)

ggsave(map_grid, filename = here("results", "map_grid.png"))

map_grid1 <- grid.arrange(rasterGrob(map_fall), rasterGrob(map_winter), ncol = 2)
ggsave(map_grid1, filename = here("results", "map_grid1.png"))
map_grid2 <- grid.arrange(rasterGrob(map_spring), rasterGrob(map_summer), ncol = 2)
ggsave(map_grid2, filename = here("results", "map_grid2.png"))
```