### 2022-03-30 code/analysis_code/results.Rmd
adding results
volunteer data and CV

---
title: "results"
output: html_document
---

#load packages and data
```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)

mp <- readRDS(here("data", "processed_data", "mpwrfcombo.rds"))
sampledata <- readRDS(here("data", "processed_data", "fullsampledata.rds"))
```

# find CV for first three sample dates (within subject A -> B)
```{r}
sample_three <- sampledata %>% filter(!date == "july_21") # remove July samples (counted by volunteers, not by self)

sample <- sample_three %>% select(a_count, b_count)
matrix_sample <- data.matrix(sample)

sample <- sample %>% mutate(cv = (Rfast::rowcvs(matrix_sample, ln = FALSE)*100))

mean(sample$cv)
median(sample$cv)
```

Woah - CVs for samples counted by me personally are averaging 20 - 26 for CV%. This is quite different from the volunteer counts. What does this say about the precision of volunteer counted-data?

# CV vs MP concentration
```{r}
# this was performed in the Barrows paper - looks at how variation correlates with concentration - previous findings showed that high-concentration sites had less variability among replicates

# create dataframe for all sites with CV values and concentration values
# average A and C and B and D counts so that they can be compared to the other sample dates for variation
avg_cd <- sampledata %>% filter(date == "july_21") %>% mutate(a_count = ((a_count+c_count)/2)) %>% mutate(b_count = ((b_count+d_count/2)))

# combine averaged july data with other three seasons of data
all_counts_cv <- full_join(avg_cd, sample_three)

# keep only useful variables
all_counts_part <- all_counts_cv %>% select(a_count, b_count, particles_l, date)
all_counts_cv <- all_counts_cv %>% select(a_count, b_count) # don't include particles/L so it won't be calculated with CV

# create matrix
all_counts_matrix <- data.matrix(all_counts_cv)
# mutate new variable to add CV
all_counts_cv <- all_counts_cv %>% mutate(cv = (Rfast::rowcvs(all_counts_matrix, ln = FALSE)*100))
# add MP concentration back in using join
all_counts_cv <- left_join(all_counts_cv, all_counts_part)
```
# plot MP concentration vs CV
```{r}
ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() + geom_smooth(method = lm) + labs(title = "Coefficient of Variance vs. Microplastic Concentration ", x = "Particles/L", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') # use log scale to better visualize all points
```

Wonder if I can make this graph and have point colors correlate to sample dates. That would be interesting
# Adding color to graph
```{r}
ggplot(all_counts_cv, aes(x = particles_l, y = cv, color = date)) + geom_point() + labs(title = "Coefficient of Variance vs. Microplastic Concentration ", x = "Particles/L", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') # use log scale to better visualize all points
```

There appears to be a negative correlation between CV and microplastic concentration, which is consistent with what's reported in the Barrows paper

## Next, let's do a Kruskal Wallis test (basically a non-parametic one-way ANOVA, that considers more than two groups)
```{r}
kruskal.test(particles_l ~ date, data = all_counts_cv)
```

P=value less than 0.05 - we conclude that there IS a significant difference between groups (though we don't yet know between what groups). From here, we may use the pairwise.wilcox.test() to calculate pairwise comparison
# pairwise wilcox test to see differences between groups
```{r}
pairwise.wilcox.test(all_counts_cv$particles_l, all_counts_cv$date,
                 p.adjust.method = "BH")
```

The only significant difference is between April and February sample dates. (in Barrows, March had the highest concentration)

Could perform regression analysis to compare original samples and corresponding duplicates over 1 year study period (using this language from the Barrows paper)
(or does this mean between A and B duplicates?) Let's try
# linear regression for duplicates over study period
```{r}
#linear regression model specification set up
lm_mod <- linear_reg() %>% set_engine("lm")
#estimating/training linear model
lm_fit_dup <- lm_mod %>% fit(a_count ~ b_count, data = all_counts_cv)
#save table
lm_fit_dup_table <- tidy(lm_fit_dup)
#view summary of fit
(lm_fit_dup)
summary(lm_fit_dup)
#save file
saveRDS(lm_fit_dup_table, file = here("results", "lm_fit_dup_table.rds"))

model = lm(a_count ~ b_count, data = all_counts_cv)
summary(model)
```

p-value is <0.05 and multiple r-squared value is 0.7824
p-value indicates that the model reflects a signficant relationship
r-squared is "degree to which the data is explained by the model" --> 78% of variability is explained by the model

```{r}
#visualize linear model
lm_dup <- ggplot(all_counts_cv, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "MP Concentration Duplicate Counts", x = "A Count", y = "B Count")
lm_dup
```

It appears that the duplicate counts are pretty well correlated.

# want to test for significant difference within a group to see if sites are signficantly different/if the distribution of MP concentrations is signficant
```{r}
kruskal.test(particles_l ~ site, data = sampledata)
```

Seen in Barrows paper - other researchers found concentrations at 70 - 100 particles/L

Try to make plot to compare sites over seasons
```{r}
sampledata %>% filter(date == "nov_20") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')
```

```{r}
# feb 2021
sampledata %>% filter(date == "feb_21") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')

# april 2021
sampledata %>% filter(date == "apr_21") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')

# july 2021
sampledata %>% filter(date == "july_21") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')
```

It could be interesting to have a table listing all of the site, and the concentration for each sample date in columns to the side. this might be easier to interpret than a graph for this type of data. 
```{r}
table <- sampledata %>% select(particles_l, site, date)

table %>% rowwise()
reshape(sampledata, idvar = "site", timevar = "date", direction = "wide")
```
```{r}
table$date <- factor(table$date , levels=c("nov_20", "feb_21", "apr_21", "july_21")) #reorder dates to appear chronologically
# transform to wide data
data_wide <- spread(table, date, particles_l)
data_wide #display new data
# display decimal places upto 1
data_wide$nov_20<-format(round(data_wide$nov_20,1),nsmall=1)
data_wide$feb_21<-format(round(data_wide$feb_21,1),nsmall=1)
data_wide$apr_21<-format(round(data_wide$apr_21,1),nsmall=1)
data_wide$july_21<-format(round(data_wide$july_21,1),nsmall=1)
```

goals: 
- look at differences in average watershed concentration over time

- test for difference between watershed groups
```{r}
#difference between watershed groups
kruskal.test(particles_l ~ watershed, data = sampledata)
```

P-value is greater than 0.05, there is NOT a signficant difference between watershed concentrations

#avg watershed concentration over time
```{r}
watershed_time <- group_by(sampledata, watershed, date) %>%
             summarise(particles_l = mean(particles_l, na.rm = TRUE))
kruskal.test(particles_l ~ date, data = watershed_time)

barber_creek <- watershed_time %>% filter(watershed == "Barber Creek")
kruskal.test(particles_l ~ date, barber_creek)

hunnicutt_creek <- watershed_time %>% filter(watershed == "Hunnicutt Creek")
kruskal.test(particles_l ~ date, hunnicutt_creek)

north_oconee_river <- watershed_time %>% filter(watershed == "North Oconee River")
kruskal.test(particles_l ~ date, north_oconee_river)

```