### 2022-04-08 code/analysis_code/particle_type.Rmd
new figures

---
title: "particle_type"
output: html_document
---

```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)
library(readxl) #for loading Excel files

type <- read_excel(here("data", "raw_data", "copy_part.xlsx"))
```

# cleaning data
```{r}
# this data set is a litte larger than we need. let's remove some unneccesary variables
a_and_b <- type %>% select("fragment", 
                           "fiber", 
                           "bead", 
                           "foam", 
                           "pellet",
                           "film",
                           "fragment_b", 
                           "fiber_b", 
                           "bead_b", 
                           "foam_b", 
                           "pellet_b", 
                           "film_b")

c_and_d <- type %>% select("fragment_c", 
                           "fiber_c", 
                           "bead_c", 
                           "foam_c", 
                           "pellet_c",
                           "film_c",
                           "fragment_d", 
                           "fiber_d", 
                           "bead_d", 
                           "foam_d", 
                           "pellet_d", 
                           "film_d")
summary(a_and_b)
```

```{r}
#summary for a and b counts
colSums(a_and_b, na.rm = TRUE)

#summary for c and d counts
colSums(c_and_d, na.rm = TRUE)
```

```{r}
part <- type %>% select("fragment", 
                           "fiber", 
                           "bead", 
                           "foam", 
                           "pellet",
                           "film",
                           "fragment_b", 
                           "fiber_b", 
                           "bead_b", 
                           "foam_b", 
                           "pellet_b", 
                           "film_b", 
                        "fragment_c", 
                           "fiber_c", 
                           "bead_c", 
                           "foam_c", 
                           "pellet_c",
                           "film_c",
                           "fragment_d", 
                           "fiber_d", 
                           "bead_d", 
                           "foam_d", 
                           "pellet_d", 
                           "film_d")
part[is.na(part)] <- 0 # replace NA with 0

part <- part %>% mutate(fragment_avg_a = ((fragment + fragment_c)/2)) 
part <- part %>% mutate(fragment_avg_b = ((fragment_b + fragment_d)/2)) 
part <- part %>% mutate(fiber_avg_a = ((fiber + fiber_c)/2)) 
part <- part %>% mutate(fiber_avg_b = ((fiber_b + fiber_d)/2))
part <- part %>% mutate(bead_avg_a = ((bead + bead_c)/2))  
part <- part %>% mutate(bead_avg_b = ((bead_b + bead_d)/2))     
part <- part %>% mutate(foam_avg_a = ((foam + foam_c)/2)) 
part <- part %>% mutate(foam_avg_b = ((foam_b + foam_d)/2)) 
part <- part %>% mutate(pellet_avg_a = ((pellet + pellet_c)/2))
part <- part %>% mutate(pellet_avg_b = ((pellet_b + pellet_d)/2)) 
part <- part %>% mutate(film_avg_a = ((film + film_c)/2))
part <- part %>% mutate(film_avg_b = ((film_b + film_d)/2))

sum <- part %>% select(25:36)
colSums(sum)

sum <- sum %>% mutate(fragment = fragment_avg_a + fragment_avg_b, 
                      fiber = fiber_avg_a + fiber_avg_b, 
                      bead = bead_avg_a + bead_avg_b, 
                      foam = foam_avg_a + foam_avg_b, 
                      pellet = pellet_avg_a + pellet_avg_b, 
                      film = film_avg_a + film_avg_b) %>% select(13:18)

colSums(sum)
# this gives us a small dataframe with just the total for each particle type
part_sum <- enframe(colSums(sum))
```

Now, we have a breakdown of total particle type (added between A and B filters). A and B counts were determined by averaging the duplicate volunteer counts performed on singular filters.

Let's see how we can visualize this breakdown. Remember that this is just for the July sample date, not for all 4 sample dates (we are looking at a subset).

```{r}
part_sum %>% ggplot(aes(x = name, y = value, fill = name)) + 
  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5, position = position_dodge(width=0.9))) + ylim(0, 350) 
```

Perhaps we could add error bars, significant difference indications in this graph.