### 2022-04-12 code/analysis_code/results.Rmd

---
title: "results"
output: html_document
---

#load packages and data
```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidyverse)
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)
library(gridExtra) #for savings images in grids
library(gridExtra) #for savings images in grids
library(ggpubr) #adding stats to figures

mp <- readRDS(here("data", "processed_data", "mpwrfcombo.rds"))
sampledata <- readRDS(here("data", "processed_data", "fullsampledata.rds"))
```

# find CV for first three sample dates (within subject A -> B)
```{r}
sample_three <- sampledata %>% filter(!date == "july_21") # remove July samples (counted by volunteers, not by self)

sample <- sample_three %>% select(a_count, b_count)
matrix_sample <- data.matrix(sample)

sample <- sample %>% mutate(cv = (Rfast::rowcvs(matrix_sample, ln = FALSE)*100))

mean(sample$cv)
median(sample$cv)
```

Woah - CVs for samples counted by me personally are averaging 20 - 26 for CV%. This is quite different from the volunteer counts. What does this say about the precision of volunteer counted-data?

# CV vs MP concentration
```{r}
# this was performed in the Barrows paper - looks at how variation correlates with concentration - previous findings showed that high-concentration sites had less variability among replicates

# create dataframe for all sites with CV values and concentration values
# average A and C and B and D counts so that they can be compared to the other sample dates for variation
avg_cd <- sampledata %>% filter(date == "july_21") %>% mutate(a_count = ((a_count+c_count)/2)) %>% mutate(b_count = ((b_count+d_count/2)))

# combine averaged july data with other three seasons of data
all_counts_cv <- full_join(avg_cd, sample_three)

# keep only useful variables
all_counts_part <- all_counts_cv %>% select(a_count, b_count, particles_l, date)
all_counts_cv <- all_counts_cv %>% select(a_count, b_count) # don't include particles/L so it won't be calculated with CV

# create matrix
all_counts_matrix <- data.matrix(all_counts_cv)
# mutate new variable to add CV
all_counts_cv <- all_counts_cv %>% mutate(cv = (Rfast::rowcvs(all_counts_matrix, ln = FALSE)*100))
# add MP concentration back in using join
all_counts_cv <- left_join(all_counts_cv, all_counts_part)
```
# plot MP concentration vs CV
```{r}
ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() + geom_smooth(method = lm) + labs(title = "Coefficient of Variance vs. Microplastic Concentration ", x = "Particles/L", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') # use log scale to better visualize all points
```

#adding stats to graph
```{r}
# with line of best fit
ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() + geom_smooth(method = lm) + labs(x = "Particles per Liter", y = "Coefficient of Variance %") +
  stat_cor(method = "spearman") + scale_x_continuous(trans='log10') + theme_bw() # use log scale to better visualize all points

#no line of best fit
ggplot(all_counts_cv, aes(x = particles_l, y = cv)) + geom_point() + labs(x = "Particles per Liter", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') + theme_bw() # use log scale to better visualize all points
```

Wonder if I can make this graph and have point colors correlate to sample dates. That would be interesting
# Adding color to graph
```{r}
ggplot(all_counts_cv, aes(x = particles_l, y = cv, color = date)) + geom_point() + labs(title = "Coefficient of Variance vs. Microplastic Concentration ", x = "Particles/L", y = "Coefficient of Variance %") + scale_x_continuous(trans='log10') # use log scale to better visualize all points
```

There appears to be a negative correlation between CV and microplastic concentration, which is consistent with what's reported in the Barrows paper

## Next, let's do a Kruskal Wallis test (basically a non-parametic one-way ANOVA, that considers more than two groups)
```{r}
kruskal.test(particles_l ~ date, data = all_counts_cv)
```

P=value less than 0.05 - we conclude that there IS a significant difference between groups (though we don't yet know between what groups). From here, we may use the pairwise.wilcox.test() to calculate pairwise comparison
# pairwise wilcox test to see differences between groups
```{r}
pairwise.wilcox.test(all_counts_cv$particles_l, all_counts_cv$date,
                 p.adjust.method = "BH")

compare_means(particles_l ~ date, data = all_counts_cv, p.adjust.method = "BH")
```

The only significant difference is between April and February sample dates. (in Barrows, March had the highest concentration)

Could perform regression analysis to compare original samples and corresponding duplicates over 1 year study period (using this language from the Barrows paper)
(or does this mean between A and B duplicates?) Let's try
# linear regression for duplicates over study period
```{r}
#linear regression model specification set up
lm_mod <- linear_reg() %>% set_engine("lm")
#estimating/training linear model
lm_fit_dup <- lm_mod %>% fit(a_count ~ b_count, data = all_counts_cv)
#save table
lm_fit_dup_table <- tidy(lm_fit_dup)
#view summary of fit
(lm_fit_dup)
summary(lm_fit_dup)
#save file
saveRDS(lm_fit_dup_table, file = here("results", "lm_fit_dup_table.rds"))

model = lm(a_count ~ b_count, data = all_counts_cv)
summary(model)
```

p-value is <0.05 and multiple r-squared value is 0.7824
p-value indicates that the model reflects a signficant relationship
r-squared is "degree to which the data is explained by the model" --> 78% of variability is explained by the model

```{r}
#visualize linear model
lm_dup <- ggplot(all_counts_cv, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "MP Concentration Duplicate Counts", x = "A Count", y = "B Count") + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') + scale_x_log10() + scale_y_log10()
lm_dup
```

It appears that the duplicate counts are pretty well correlated.

We could use this same type of graph to display the comparative CV for my counts vs volunteer counts by doing AvsB for sample dates 1-3 and then AvsB for sample date 4 or AvsC and BvsD

#visualizing CV
```{r}
#visualize linear model
#subset data
all_counts_me <- all_counts_cv %>% filter(!date == "july_21") 
#plot
lm_me <- ggplot(all_counts_me, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Duplicate Counts Nov, Apr, July", x = "A Count", y = "B Count") + xlim(0,25) + ylim(0,25) + geom_abline(intercept = 0, slope = 1, colour='#E41A1C')
lm_me #show plot

#subset data
all_counts_vol <- all_counts_cv %>% filter(date == "july_21")
#plot
lm_vol <- ggplot(all_counts_vol, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Volunteer Counts July", x = "A Count", y = "B Count") + xlim(0,25) + ylim(0,25) + geom_abline(intercept = 0, slope = 1, colour='#E41A1C')
lm_vol #show plot
```

# CV for each sample date
```{r}
#nov
nov <- all_counts_cv %>% filter(date == "nov_20")
#plot
lm_nov <- ggplot(nov, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Nov", x = "A Count", y = "B Count") + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') + expand_limits(x = 0, y = 0) 
lm_nov #show plot

#feb
feb <- all_counts_cv %>% filter(date == "feb_21")
#plot
lm_feb <- ggplot(feb, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Feb", x = "A Count", y = "B Count") + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') 
lm_feb #show plot

#april
april <- all_counts_cv %>% filter(date == "apr_21")
#plot
lm_apr <- ggplot(april, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "April", x = "A Count", y = "B Count") + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') 
lm_apr #show plot

#july
all_counts_vol <- all_counts_cv %>% filter(date == "july_21")
#plot
lm_vol <- ggplot(all_counts_vol, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "July", x = "A Count", y = "B Count") + ylim(0,25) + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') + xlim(3,25) + ylim(3,30)
lm_vol #show plot

bivar_grid <- grid.arrange(lm_nov, lm_feb, lm_apr, lm_vol, ncol = 2)
bivar_grid #show grid
```

# want to test for significant difference within a group to see if sites are signficantly different/if the distribution of MP concentrations is signficant
```{r}
kruskal.test(particles_l ~ site, data = sampledata)
```

Seen in Barrows paper - other researchers found concentrations at 70 - 100 particles/L

Try to make plot to compare sites over seasons
```{r}
sampledata %>% filter(date == "nov_20") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')
```

```{r}
# feb 2021
sampledata %>% filter(date == "feb_21") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')

# april 2021
sampledata %>% filter(date == "apr_21") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')

# july 2021
sampledata %>% filter(date == "july_21") %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_point()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) + scale_y_continuous(trans='log10')
```

It could be interesting to have a table listing all of the site, and the concentration for each sample date in columns to the side. this might be easier to interpret than a graph for this type of data. 
```{r}
table <- sampledata %>% select(particles_l, site, date)

table %>% rowwise()
reshape(sampledata, idvar = "site", timevar = "date", direction = "wide")
```
```{r}
table$date <- factor(table$date , levels=c("nov_20", "feb_21", "apr_21", "july_21")) #reorder dates to appear chronologically
# transform to wide data
data_wide <- spread(table, date, particles_l)
data_wide #display new data
# display decimal places upto 1
data_wide$nov_20<-format(round(data_wide$nov_20,1),nsmall=1)
data_wide$feb_21<-format(round(data_wide$feb_21,1),nsmall=1)
data_wide$apr_21<-format(round(data_wide$apr_21,1),nsmall=1)
data_wide$july_21<-format(round(data_wide$july_21,1),nsmall=1)
```

goals: 
- look at differences in average watershed concentration over time

- test for difference between watershed groups
```{r}
#difference between watershed groups
kruskal.test(particles_l ~ watershed, data = sampledata)
```

P-value is greater than 0.05, there is NOT a signficant difference between watershed concentrations

#avg watershed concentration over time
```{r}
watershed_time <- group_by(sampledata, watershed, date) %>%
             summarise(particles_l = mean(particles_l, na.rm = TRUE))
kruskal.test(particles_l ~ date, data = watershed_time)

barber_creek <- watershed_time %>% filter(watershed == "Barber Creek")
kruskal.test(particles_l ~ date, barber_creek)

hunnicutt_creek <- watershed_time %>% filter(watershed == "Hunnicutt Creek")
kruskal.test(particles_l ~ date, hunnicutt_creek)

north_oconee_river <- watershed_time %>% filter(watershed == "North Oconee River")
kruskal.test(particles_l ~ date, north_oconee_river)

```
# main rivers vs tributaries
```{r}
river <- sampledata %>% filter(watershed == "North Oconee River"|watershed == "Middle Oconee River"|watershed == "Oconee River")

trib <- sampledata %>% filter(!watershed == "North Oconee River", !watershed == "Middle Oconee River", !watershed == "Oconee River")

river <- river %>% mutate(body = "river")
trib <- trib %>% mutate(body = "trib")

#rejoin
body <- full_join(river, trib)

#rename
body$body <- fct_recode(body$body,
                                 "Main Rivers" = "river",
                                 "Tributaries" = "trib")
pv <- body %>% summarize(p.value = round(kruskal.test(particles_l ~ body)$p.value, digits = 3))

#plot
body %>% ggplot(aes(x = body, y = particles_l, color = )) + 
  geom_boxplot()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(x = "Water Body", y = "Particles per Liter") + ylim(0,500) + scale_y_log10() +
    geom_text(data=pv, aes(x=1.5, y=900, label=paste0("Kruskal-Wallis\n p=",p.value))) + theme_bw()

#kruskal wallis/mann whitney test - nonparametric test
kruskal.test(particles_l ~ body, data = body)
wilcox.test(particles_l ~ body, data = body)

summary(river)
summary(trib)
```

P=value less than 0.05 - we conclude that there IS a significant difference between groups. Let's do a pairwise.wilcox.test() to double check. 
# pairwise wilcox test to see differences between groups
```{r}
pairwise.wilcox.test(body$particles_l, body$body,
                 p.adjust.method = "BH")
```

This tells us that there is a significant difference between the MP concentrations in rivers and tributaries! Cool! We can try to tweak this boxplot figure to display those statistics, perhaps. 
# figure with stats for seasons
```{r}
compare_means(particles_l ~ date,  data = body)
my_comparisons <- list(c("nov_20", "feb_21"), c("nov_20", "apr_21"), c("nov_20", "july_21"), c("feb_21", "apr_21"), c("feb_21", "july_21"), c("apr_21", "july_21"))

fig_season <- body %>% ggplot(aes(x = date, y = particles_l, group = date)) + 
  geom_boxplot() + scale_color_discrete(name = "Water Body")+ 
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(x = "Sample Date", y = "Particles per Liter") + scale_x_discrete(labels=c("Fall 2020",
                            "Winter 2021",
                            "Spring 2021",
                            "Summer 2021"))+ theme_bw() + stat_compare_means(comparison=my_comparisons, p.adjust.method = "BH") + stat_compare_means(label.y = 5) + scale_y_log10()
plot(fig_season)
```
#arrange figures
```{r}
#map
#Load Libraries 
library(tidyverse)
library(viridis)
library(ggpubr)
#Load data
sites = readRDS(here::here("data","processed_data","mpwrfcombo.rds"))
sites$date = as.factor(sites$date)
sites$date = recode(sites$date, nov_20 = "Nov 2020", feb_21 = "Feb 2021", apr_21 = "April 2021", july_21 = "July 2021")
sites$date = ordered(sites$date, levels = c("Nov 2020", "Feb 2021", "April 2021", "July 2021"))
sites %>% ggplot(aes(x = date, y = site, fill= particles_l)) + 
  geom_tile() + 
  scale_fill_viridis() + 
  theme_bw() +
  ylab("Site") + 
  xlab("Collection Period") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = date, y = site, label = round(particles_l))) +
  labs(fill = "Particles per Liter")  
  #ggtitle("Seasonality of Microplastics Concentration")
sites %>% ggplot(aes(x = date, y = site, fill= log10(particles_l))) + 
  geom_tile() + 
  scale_fill_viridis() + 
  theme_bw() +
  ylab("Site") + 
  xlab("") + 
  scale_x_discrete(position = "top") +
  #theme(axis.text.x = element_text(angle = 45)) +
  geom_text(aes(x = date, y = site, label = round(log10(particles_l), 2))) +
  labs(fill = "Log10(Particles/Liter)") + 
  theme(legend.position = "bottom")
  #ggtitle("Seasonality of Microplastics Concentration")
hist_all = sites %>% ggplot(aes(x = particles_l)) + 
  geom_histogram() + 
  theme_bw() + 
  #facet_wrap(~date) +
  xlab("Particles per Liter") + 
  ylab("No. of Observations") +
  theme(axis.text.y=element_text(size=12))
hist_season = sites %>% ggplot(aes(x = particles_l)) + 
  geom_histogram() + 
  theme_bw() + 
  facet_wrap(~date) +
  xlab("Particles per Liter") + 
  ylab("No. of Observations") +
  theme(axis.text.y=element_text(size=12))
fig2 = ggarrange(hist_all, hist_season, ncol = 2)
plot(fig2)
sites %>% group_by(date) %>% tally()

#particle type = fig_part_type
#particle color = fig_part_color
quad1 <- ggarrange(hist_all, hist_season, fig_part_type, fig_part_color, nrow = 2, ncol = 2, labels = "AUTO", vjust = 1) 
#%>% ggexport(filename = here("results","manuscript_figures", "quad_1.png"))
plot(quad_1)

ggsave(filename = here("results","manuscript_figures", "quad_1.png")) #export resolution than ggexport
```